---
title: "Development and validation of the FAI scale"
subtitle: "Diagnosis Category as Criterion Variable"
author: Corrado Caudek 
date: today
format:
  html:
    embed-resources: true
    standalone: true
    toc: true
    code-fold: show
    code-copy: true
    code-line-numbers: true
editor_options: 
  chunk_output_type: console
bibliography: refs.bib
execute:
  warning: false
  echo: true
---

## Background

This study aims to develop a psychological scale to differentiate between family caregivers of children with chronic illnesses based on their coping abilities. This scale will be used to identify caregivers who may benefit from additional support or interventions.

## Objective

In this analysis, the diagnosis category (i.e., Lieve, Grave, Moderata, Molto grave) will be used as the criterion variable in a predictive model. The purpose is to evaluate the classification ability of the FAI scales. 

## Load packages and options

```{r}
if (!requireNamespace("pacman")) install.packages("pacman")

pacman::p_load(
  here, tidyverse, TAM, mirt, lavaan, mokken, psych, semTools, MASS, brant, 
  ids, pROC, nnet, caret, ROSE, DMwR, recipes, themis, randomForest, reshape2,
  VGAM, mgcv
)

# Disable significance stars globally
options(show.signif.stars = FALSE)
```

## Import the Data

```{r}
# Import RedCap data ------------------------------------------------------

setwd("~/_repositories/meyer_fai/scripts/redcap")

# This will load the necessary libreries
source(here::here("scripts", "redcap", "_targets.R"))

# Read cleaned RedCap data
temp <- tar_read(fai_clean_complete)
# Some cases are empty and will be ignored: 12 18 76 81 152 182 193 202
temp1 <- temp[-c(12, 18, 76, 81, 152, 182, 193, 202), ]
# Remove outlier found with the performance package
fai_clean_complete <- temp1[-92, ]
rm(temp, temp1)

fai <- fai_clean_complete |> 
  dplyr::select(
    fai_child_charact, fai_caregiving, fai_cure, fai_intrapsych, 
    fai_support, fai_hyper, quanti_fratelli, sesso, nome_malattia,
    eta_pz, legge_104
  ) |> 
   dplyr::rename(
     fai_child_char = fai_child_charact,
     fai_caregiving = fai_caregiving,
     fai_percz_cure = fai_cure,
     fai_intra_psyc = fai_intrapsych,
     fai_soc_fam_sp = fai_support,
     fai_hyper_prot = fai_hyper
     #eta_madre = erta_madre
   ) 
```

Determine severity level:

```{r}
fai <- fai |> 
  mutate(
    nome_malattia_raggruppato = case_when(
      # Molto grave: malattie con alto rischio di morte o che 
      # richiedono trattamenti intensivi
      nome_malattia %in% c(
        "glioma", "LLA", "linfoma anaplastico", "Leucemia mieloide", 
        "fibrosi cistica", "Pinealoblastoma", 
        "tumore glioneurale leptomeningeale diffuso"
        ) ~ "Molto grave",
      
      # Grave: malattie croniche, con impatto significativo sulla 
      # qualità della vita
      nome_malattia %in% c(
        "Epilessia", "sindrome di charge", "Distrofia muscolare",
        "Neurofibromatosi", "diabete di tipo 1", "artrite romatoide",
        "RCU", "emofilia"
        ) ~ "Grave",
      
      # Moderata: malattie con impatto moderato, trattabili ma non 
      # trascurabili
      nome_malattia %in% c(
        "asma", "Emicrania", "dermatite atopica", "ipospadia", 
        "ipotonia muscolare", "Blefaro chetato congiuntivite"
        ) ~ "Moderata",
      
      # Lieve: condizioni gestibili con trattamenti standard e 
      # basso impatto funzionale
      TRUE ~ "Lieve"
    )
  )

fai$nome_malattia_raggruppato <- factor(fai$nome_malattia_raggruppato)
fai$nome_malattia_raggruppato <- 
  relevel(fai$nome_malattia_raggruppato, ref = "Lieve")

fai$nome_malattia <- NULL

# Confirm levels of the categorical variable
levels(fai$nome_malattia_raggruppato)
```

## Check for Missing Values

```{r}
# Check for missing values in the dataset
# This ensures there are no NA values that could interfere with model fitting 
# or predictions
sum(is.na(fai))
```

## Inspecting the Dataset

Display class distribution.

```{r}
# Display the distribution of the outcome variable to understand class imbalance
table(fai$nome_malattia_raggruppato)
```

Visualize class proportions.

```{r}
# Visualize the class distribution of the outcome variable
ggplot(data.frame(fai$nome_malattia_raggruppato), aes(x = fai$nome_malattia_raggruppato)) +
  geom_bar() +
  labs(title = "Class Distribution of Disease Severity",
       x = "Disease Severity",
       y = "Count")
```

## Addressing Class Imbalance

```{r}
set.seed(123)

# Use recipe to balance the dataset by upsampling minority classes
recipe_obj <- recipe(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + 
    fai_hyper_prot +
    eta_pz + quanti_fratelli, 
  data = fai
) %>%
  step_upsample(nome_malattia_raggruppato, over_ratio = 1)

# Prepare the balanced dataset
balanced_fai <- prep(recipe_obj) %>% juice()

# Verify that the classes are now balanced
table(balanced_fai$nome_malattia_raggruppato)
```

## Fitting the Multinomial Logistic Regression Model

A multinomial model was used because 

```{r}
set.seed(123)

# Fit a multinomial logistic regression model
# The predictors are selected based on their relevance
fm <- multinom(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving + 
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot + 
    eta_pz + quanti_fratelli,
  data = balanced_fai
)
```

Display the model summary.

```{r}
# Summarize the model to view coefficients for each class
summary(fm)
```

Calculate z-values and p-values for model coefficients.

```{r}
# Calculate z-values and p-values for model coefficients
z_values <- summary(fm)$coefficients / summary(fm)$standard.errors
p_values <- (1 - pnorm(abs(z_values), 0, 1)) * 2

# Display p-values to assess the significance of predictors
print(p_values)
```

## Evaluating Predictive Power

Predictions and Confusion Matrix

```{r}
# Predict class labels on the training data
predicted <- predict(fm, balanced_fai)

# Generate and display a confusion matrix
confusion_matrix <- table(Predicted = predicted, Actual = balanced_fai$nome_malattia_raggruppato)
print(confusion_matrix)

# Calculate and display the accuracy of the model
accuracy <- mean(predicted == balanced_fai$nome_malattia_raggruppato)
print(paste("Accuracy:", round(accuracy, 2)))
```

Compare accuracy to baseline (random guessing).

```{r}
# Calculate the baseline accuracy (random guessing)
baseline_accuracy <- 1 / nlevels(balanced_fai$nome_malattia_raggruppato)
print(paste("Baseline Accuracy:", round(baseline_accuracy, 2)))
```

## Cross-Validation

```{r}
#| output: false
set.seed(123)

# Perform 10-fold cross-validation to evaluate the model's generalizability
train_control <- trainControl(method = "cv", number = 10)
cv_model <- train(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    eta_pz + quanti_fratelli,
  data = balanced_fai,
  method = "multinom",
  trControl = train_control
)
```

```{r}
# Print cross-validation results
print(cv_model)
```


## Relative Importance of Predictors

Display Variable Importance.

```{r}
set.seed(123)

# Assess variable importance using caret
importance_model <- varImp(fm, scale = FALSE)
print(importance_model)
```

Visualize relative importance.

```{r}
# Extract and reshape coefficient data for plotting
coef_data <- as.data.frame(t(summary(fm)$coefficients))
coef_data$Variable <- rownames(coef_data)

# Reshape the coefficients for visualization
coef_long <- reshape2::melt(coef_data, id.vars = "Variable")

# Plot absolute coefficient values for predictors
ggplot(coef_long, aes(x = reorder(Variable, abs(value)), y = abs(value), fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(title = "Absolute Coefficients for Predictors",
       x = "Predictor",
       y = "Coefficient Magnitude",
       fill = "Outcome Category")

```

## Visualizing Predicted Probabilities

```{r}
# Create a new dataset to visualize the effect of fai_child_char
new_data <- data.frame(
  fai_child_char = seq(min(fai$fai_child_char), max(fai$fai_child_char), length.out = 100),
  fai_caregiving = mean(fai$fai_caregiving, na.rm = TRUE),
  fai_percz_cure = mean(fai$fai_percz_cure, na.rm = TRUE),
  fai_intra_psyc = mean(fai$fai_intra_psyc, na.rm = TRUE),
  fai_soc_fam_sp = mean(fai$fai_soc_fam_sp, na.rm = TRUE),
  fai_hyper_prot = mean(fai$fai_hyper_prot, na.rm = TRUE),
  eta_pz = mean(fai$eta_pz, na.rm = TRUE),
  quanti_fratelli = mean(fai$quanti_fratelli, na.rm = TRUE)
)

# Predict probabilities for the new data
probs <- predict(fm, newdata = new_data, type = "probs")
new_data <- cbind(new_data, probs)

# Reshape data for plotting
plot_data <- melt(new_data, id.vars = "fai_child_char", variable.name = "Disease_Severity", value.name = "Probability")

# Plot the predicted probabilities
ggplot(plot_data, aes(x = fai_child_char, y = Probability, color = Disease_Severity)) +
  geom_line() +
  labs(title = "Predicted Probabilities of Disease Severity",
       x = "FAI Child Characteristics",
       y = "Predicted Probability")
```


## Variable Importance from Random Forest

```{r}
set.seed(123)
rf_model <- randomForest(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    eta_pz + quanti_fratelli,
  data = balanced_fai,
  ntree = 500,
  importance = TRUE
)

# Plot variable importance
varImpPlot(rf_model)
```

## Ensemble Predictions

Using ensemble methods like combining predictions from Random Forest and multinomial logistic regression (majority vote) can improve predictive performance. 

```{r}
# Combine predictions (majority vote)
ensemble_preds <- apply(cbind(predict(fm, balanced_fai), predict(rf_model, balanced_fai)), 1, function(x) {
  names(which.max(table(x)))
})

# Evaluate ensemble predictions
conf_matrix <- confusionMatrix(
  data = factor(ensemble_preds, levels = levels(balanced_fai$nome_malattia_raggruppato)),
  reference = balanced_fai$nome_malattia_raggruppato
)
print(conf_matrix)
```

## Ordinal Logistic Regression

```{r}
ord_model <- polr(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    eta_pz + quanti_fratelli,
  data = balanced_fai,
  method = "logistic"
)

# Model summary
summary(ord_model)
```

```{r}
# Fit a null model (intercept-only model)
null_model <- polr(
  nome_malattia_raggruppato ~ 1,
  data = balanced_fai,
  method = "logistic"
)

# Compare the models using a likelihood ratio test
lr_test <- anova(null_model, ord_model, test = "Chisq")
print(lr_test)
```

A significant p-value suggests that the predictors significantly improve the model's fit compared to the null model.

```{r}
# Extract coefficients and standard errors
coefs <- summary(ord_model)$coefficients
z_values <- coefs[, 1] / coefs[, 2]  # z = coefficient / std. error
p_values <- 2 * (1 - pnorm(abs(z_values)))  # Two-tailed p-values

# Display coefficients with significance
coef_table <- data.frame(
  Estimate = coefs[, 1],
  Std_Error = coefs[, 2],
  Z_Value = z_values,
  P_Value = p_values
)
print(coef_table)
```

Predictors with low p-values contribute significantly to predicting the ordinal outcome.

Predictive Accuracy: Concordance Index

The concordance index (C-index) measures how well the model predicts the order of the response variable. A C-index of 0.5 indicates random guessing, while values closer to 1 indicate better predictive power.

Calculate C-index

```{r}
# Predict probabilities
predicted_probs <- predict(ord_model, type = "probs")

# Calculate concordance index using Hmisc package
library(Hmisc)
c_index <- rcorr.cens(predicted_probs[, 1], as.numeric(balanced_fai$nome_malattia_raggruppato))
print(c_index)
```

```{r}
brant(ord_model)
```



The ordinal logistic regression model may not be appropriate for your data because:

The proportional odds assumption is violated.
Some predictors show highly variable effects across outcome levels.
You will need to consider alternative models or approaches to properly evaluate the predictive power of the predictors.


```{r}
partial_pom <- vglm(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    eta_pz + quanti_fratelli,
  family = cumulative(parallel = FALSE),  # Relax proportional odds
  data = balanced_fai
)

# Model summary
summary(partial_pom)
```


```{r}
rf_model <- randomForest(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    eta_pz + quanti_fratelli,
  data = balanced_fai,
  ntree = 500,
  importance = TRUE
)

# Variable importance
varImpPlot(rf_model)
```







```{r}
# Fit multinomial logistic regression
m1 <- multinom(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    sesso +
    eta_pz +
    quanti_fratelli + legge_104,
  data = fai
)
```

```{r}
# Summary of the model
summary(m1)
```

```{r}
# Calculate z-values and p-values
z_values <- summary(m1)$coefficients / summary(m1)$standard.errors
p_values <- (1 - pnorm(abs(z_values), 0, 1)) * 2

# Display p-values
print(p_values)
```

```{r}
# Predict on the training data
predicted <- predict(m1, balanced_fai)

# Confusion matrix
table(Predicted = predicted, Actual = balanced_fai$nome_malattia_raggruppato)

# Calculate accuracy
accuracy <- mean(predicted == balanced_fai$nome_malattia_raggruppato)
print(paste("Accuracy:", round(accuracy, 2)))
```


```{r}
# Example: Visualize predicted probabilities for one variable
new_data <- data.frame(
  fai_child_char = seq(min(fai$fai_child_char), max(fai$fai_child_char), length.out = 100),
  fai_caregiving = mean(fai$fai_caregiving),
  fai_percz_cure = mean(fai$fai_percz_cure),
  fai_intra_psyc = mean(fai$fai_intra_psyc),
  fai_soc_fam_sp = mean(fai$fai_soc_fam_sp),
  fai_hyper_prot = mean(fai$fai_hyper_prot)
)

# Predicted probabilities
probs <- predict(m1, newdata = new_data, type = "probs")

# Add probabilities to new_data
new_data <- cbind(new_data, probs)

# Reshape data for plotting
library(reshape2)
plot_data <- melt(new_data, id.vars = "fai_child_char", variable.name = "Disease_Severity", value.name = "Probability")

# Plot
ggplot(plot_data, aes(x = fai_child_char, y = Probability, color = Disease_Severity)) +
  geom_line() +
  labs(title = "Predicted Probabilities of Disease Severity",
       x = "FAI Child Characteristics",
       y = "Probability")
```


## Address Class Imbalance

Given the imbalance in severity levels, apply techniques to ensure the model learns effectively from minority classes.

Oversampling with Recipes
Use recipes to balance the dataset by oversampling:

```{r}
set.seed(123)

# Create an upsampling recipe
recipe_obj <- recipe(nome_malattia_raggruppato ~ ., data = fai) %>%
  step_upsample(nome_malattia_raggruppato, over_ratio = 1)

# Prepare balanced data
balanced_fai <- prep(recipe_obj) %>% juice()
table(balanced_fai$nome_malattia_raggruppato)
```


### Build Models
(a) Multinomial Logistic Regression
This model is interpretable and works well for multiclass classification.

```{r}
# Fit multinomial logistic regression
set.seed(123)
fm <- multinom(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure + fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot +
    sesso + eta_pz +
    quanti_fratelli + legge_104,
  data = balanced_fai
)

# Model summary
summary(fm)
```

```{r}
# Calculate z-values and p-values
z_values <- summary(fm)$coefficients / summary(fm)$standard.errors
p_values <- (1 - pnorm(abs(z_values), 0, 1)) * 2

# Display p-values
print(p_values)
```

```{r}
# Predict on the training data
predicted <- predict(fm, balanced_fai)

# Confusion matrix
table(Predicted = predicted, Actual = balanced_fai$nome_malattia_raggruppato)

# Calculate accuracy
accuracy <- mean(predicted == balanced_fai$nome_malattia_raggruppato)
print(paste("Accuracy:", round(accuracy, 2)))
```


### Fit Random Forest

```{r}
set.seed(123)
rf_model <- randomForest(
  nome_malattia_raggruppato ~ fai_child_char + fai_caregiving +
    fai_percz_cure +
    fai_intra_psyc + fai_soc_fam_sp + fai_hyper_prot + sesso +
    eta_pz +
    quanti_fratelli + legge_104,
  data = balanced_fai,
  ntree = 500,
  importance = TRUE
)

# Variable importance
varImpPlot(rf_model)
```

### XGBoost

```{r}
library(xgboost)

# Prepare data for XGBoost
train_matrix <- model.matrix(nome_malattia_raggruppato ~ ., data = balanced_fai)[, -1]
train_labels <- as.numeric(balanced_fai$nome_malattia_raggruppato) - 1

# Fit XGBoost
set.seed(123)
xgb_model <- xgboost(
  data = as.matrix(train_matrix),
  label = train_labels,
  objective = "multi:softmax",
  num_class = length(unique(fai$nome_malattia_raggruppato)),
  nrounds = 100,
  eta = 0.3,
  max_depth = 6,
  eval_metric = "mlogloss"
)
```


### Evaluate Model Performance
Predictions on Test Data
Use stratified train-test splitting to ensure balanced evaluation:

```{r}
# Create train-test split
set.seed(123)
train_index <- createDataPartition(
  fai$nome_malattia_raggruppato, p = 0.7, list = FALSE)
train <- fai[train_index, ]
test <- fai[-train_index, ]

# Predict and evaluate for Random Forest
test_predictions_rf <- predict(rf_model, newdata = test)
conf_matrix_rf <- confusionMatrix(
  data = test_predictions_rf,
  reference = test$nome_malattia_raggruppato
)
print(conf_matrix_rf)
```

Metrics:

- Accuracy: Percentage of correct predictions.
- Precision, Recall, F1-score: Assess performance for each class.
- Confusion Matrix: Evaluate class-specific performance.

```{r}
conf_matrix_rf$byClass  # Class-specific metrics
conf_matrix_rf$overall  # Overall metrics
```

### Cross-Validation

Use cross-validation to ensure robustness:

```{r}
set.seed(123)
train_control <- trainControl(method = "cv", number = 10)

# Train Random Forest with cross-validation
cv_model <- train(
  nome_malattia_raggruppato ~ .,
  data = fai,
  method = "rf",
  trControl = train_control
)
print(cv_model)
```


## Bibliography



```{r}
# Create the severity variable based on `nome_malattia`
fai_clean_complete <- fai_clean_complete %>%
  mutate(
    nome_malattia_raggruppato = case_when(
      # Molto grave: severe conditions with high mortality risk
      nome_malattia %in% c(
        "glioma", "LLA", "linfoma anaplastico", "Leucemia mieloide",
        "fibrosi cistica", "Pinealoblastoma",
        "tumore glioneurale leptomeningeale diffuso"
      ) ~ "Molto grave",

      # Grave: chronic illnesses significantly impacting quality of life
      nome_malattia %in% c(
        "Epilessia", "sindrome di charge", "Distrofia muscolare",
        "Neurofibromatosi", "diabete di tipo 1", "artrite romatoide",
        "RCU", "emofilia"
      ) ~ "Grave",

      # Moderata: moderate conditions that are treatable but impactful
      nome_malattia %in% c(
        "asma", "Emicrania", "dermatite atopica", "ipospadia",
        "ipotonia muscolare", "Blefaro chetato congiuntivite"
      ) ~ "Moderata",

      # Lieve: mild conditions manageable with standard treatments
      TRUE ~ "Lieve"
    )
  ) %>%
  mutate(
    nome_malattia_raggruppato = factor(nome_malattia_raggruppato, levels = c("Lieve", "Moderata", "Grave", "Molto grave"))
  ) %>%
  dplyr::select(-nome_malattia)  # Remove the original disease name variable

```

```{r}
table(fai_clean_complete$nome_malattia_raggruppato)
```

```{r}
fai_clean_complete <- fai_clean_complete %>%
  mutate(across(starts_with("fai"), scale))
```

```{r}
library(recipes)

set.seed(123)
# Upsampling recipe
recipe_obj <- recipe(
  nome_malattia_raggruppato ~ ., data = fai_clean_complete) %>%
  step_upsample(nome_malattia_raggruppato, over_ratio = 1)  # Balance classes

# Prepare the balanced dataset
balanced_data <- prep(recipe_obj) %>% juice()
table(balanced_data$nome_malattia_raggruppato)
```

```{r}
set.seed(123)
train_index <- createDataPartition(
  balanced_data$nome_malattia_raggruppato, p = 0.7, list = FALSE)
train_data <- balanced_data[train_index, ]
test_data <- balanced_data[-train_index, ]
```

```{r}
# Fit multinomial logistic regression
multinom_model <- multinom(
  nome_malattia_raggruppato ~ fai_child_charact + fai_caregiving +
    fai_cure + fai_intrapsych + fai_support + fai_hyper +
    sesso + eta_pz + quanti_fratelli + legge_104,
  data = train_data
)

# Model summary
summary(multinom_model)
```

```{r}
# Fit Random Forest
rf_model <- randomForest(
  nome_malattia_raggruppato ~
    fai_child_charact + fai_caregiving +
    fai_cure + fai_intrapsych + fai_support + sesso + eta_pz +
    quanti_fratelli + legge_104,
  data = train_data,
  ntree = 500,
  importance = TRUE
)

# Variable importance
varImpPlot(rf_model)
```

```{r}
# Predict using Random Forest
test_predictions_rf <- predict(rf_model, newdata = test_data)

# Evaluate performance
conf_matrix <- confusionMatrix(
  data = test_predictions_rf,
  reference = test_data$nome_malattia_raggruppato
)
print(conf_matrix)
```

```{r}
conf_matrix$byClass  # Sensitivity, specificity, and F1-score for each class
```

```{r}
# Standardize numeric predictors
fai_clean_complete <- fai_clean_complete %>%
  mutate(across(
    starts_with("fai"),
    scale,
    .names = "scaled_{.col}"
  ))
```


```{r}
# Create interaction terms
fai_clean_complete <- fai_clean_complete %>%
  mutate(
    fai_hyper_sesso = scaled_fai_hyper * as.numeric(sesso == "female"),
    fai_caregiving_age = scaled_fai_caregiving * eta_pz
  )
```

```{r}
table(fai_clean_complete$nome_malattia_raggruppato)
```

```{r}
set.seed(123)
train_index <- createDataPartition(
  fai_clean_complete$nome_malattia_raggruppato, p = 0.7, list = FALSE)
train_data <- fai_clean_complete[train_index, ]
test_data <- fai_clean_complete[-train_index, ]
```

```{r}
set.seed(123)
rf_model <- randomForest(
  nome_malattia_raggruppato ~ scaled_fai_child_charact +
    scaled_fai_caregiving +
    scaled_fai_cure + scaled_fai_intrapsych + scaled_fai_support +
    scaled_fai_hyper + fai_hyper_sesso + fai_caregiving_age + eta_pz + sesso,
  data = train_data,
  ntree = 500,
  importance = TRUE
)

# Variable importance
varImpPlot(rf_model)
```

```{r}
# Prepare data for XGBoost
train_matrix <- model.matrix(
  nome_malattia_raggruppato ~ . -1, data = train_data)
train_labels <- as.numeric(train_data$nome_malattia_raggruppato) - 1

set.seed(123)
xgb_model <- xgboost(
  data = train_matrix,
  label = train_labels,
  objective = "multi:softmax",
  num_class = length(unique(train_data$nome_malattia_raggruppato)),
  eval_metric = "mlogloss",
  nrounds = 100,
  eta = 0.3,
  max_depth = 6
)
```

```{r}
# Random Forest predictions
rf_preds <- predict(rf_model, newdata = test_data)

# XGBoost predictions
test_matrix <- model.matrix(nome_malattia_raggruppato ~ . -1, data = test_data)
xgb_preds <- predict(xgb_model, newdata = test_matrix)
xgb_preds <- as.factor(levels(train_data$nome_malattia_raggruppato)[xgb_preds + 1])

# Ensemble predictions (majority vote)
ensemble_preds <- apply(cbind(rf_preds, xgb_preds), 1, function(x) {
  names(which.max(table(x)))
})
```

```{r}
# Confusion matrix
conf_matrix <- confusionMatrix(
  data = factor(ensemble_preds, levels = levels(test_data$nome_malattia_raggruppato)),
  reference = test_data$nome_malattia_raggruppato
)

print(conf_matrix)
```

```{r}
# Fit ordinal logistic regression
ord_model <- polr(
  nome_malattia_raggruppato ~
    scaled_fai_child_charact +
    scaled_fai_caregiving +
    scaled_fai_cure + scaled_fai_intrapsych +
    scaled_fai_support +
    scaled_fai_hyper
  ,
  data = fai_clean_complete,
  method = "logistic"
)

# Summarize the model
summary(ord_model)
```

###################

```{r}
library(dplyr)
library(cmdstanr)

# Dati di esempio
set.seed(123)
N <- 100  # Numero di osservazioni
K <- 4    # Numero di categorie
data <- data.frame(
  X1 = rnorm(N),
  X2 = rnorm(N),
  y = sample(1:K, N, replace = TRUE, prob = c(0.7, 0.2, 0.08, 0.02)) # sbilanciati
)

# Calcolo dei pesi inversi
class_weights <- data %>%
  count(y) %>%
  mutate(weight = 1 / n)

# Unione dei pesi ai dati
data <- data %>%
  left_join(class_weights, by = "y") %>%
  rename(weight = weight)

# Preparare i dati per Stan
stan_data <- list(
  N = nrow(data),
  K = K,
  P = 2,  # Numero di covariate
  X = as.matrix(data[, c("X1", "X2")]),
  y = data$y,
  weights = data$weight
)
```


```{r}
# Specificare il percorso al file Stan
model_file <- here::here("scripts", "redcap", "multinom_logistic_reg.stan")

# Compilare il modello
model <- cmdstan_model(model_file)

# Run the model
fit <- model$sample(
  data = stan_data,
  seed = 123,
  chains = 4,
  iter_sampling = 2000,
  iter_warmup = 1000
)
```

```{r}
library(posterior)

# Extract posterior draws
posterior_draws <- as_draws_df(fit$draws())

# View parameter summaries
summary_draws <- summarise_draws(posterior_draws)
print(summary_draws)

```


